On the link between functional traits and growth rate: meta-analysis shows effects change with plant size, as predicted

### 2016-03-16 MS.tex
Small editing

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``seedling'', ``sapling'', or ``adult''. The group ``juvenile'' was largely overlapping with the three others, and so we merged this category with one or the other on a case by case basis. Using the above cut-offs, we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 7 of 103 articles were ambiguous on their ontogenetic stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck} -- were sometimes assessed as adult when reproductive. For purposes of the present paper, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. We found that results using the original categories established by the authors were similar to those using the standardized categories (see Fig. \ref{fig:figA6}). Thus in the main text, we report only results using the standardized categories.

@@

The dataset included studies differing in the growth measurement used (relative or absolute), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. These differences provide good criteria for assessing the quality of the data for addressing our hypotheses. We adopted a three-stage approach where at first we were conservative by focusing exclusively on relative growth rate, across studies that included \underline{\textgreater} 10 species, under unstressed conditions, and where the trait and growth were measured at the same stage (expect for maximum height and seed mass, where a particular stage is by definition required to measure the trait). We then compared these results with those from the subset of the dataset using relative growth rate, but with all other criteria relaxed. Finally, we compared results from both relative and absolute growth rates, from the entire dataset.

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero).

For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistent effect in the coefficient reported for each stage across studies.

As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and Aarea, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size.

For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.

For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

For five prominent traits, the patterns observed were consistent with specific predictions from a mechanistic growth model, although there were some limitations arising from unequal numbers of observation across ontogenetic stages. This suggests the mechanisms through which traits affect growth are correctly represented in the model. For SLA the meta-analysis strongly confirmed the predicted influence of plant size on the trait-growth correlation. For wood density, Aarea, seed mass and maximum height results were consistent with predictions, but data was limited and additional studies would be helpful for these trait-growth-stage relationships.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1}
\caption{Number of recorded correlations between trait and growth rate in our dataset, by stage and trait. Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic or maximum height, Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea = net mass assimilation rate per unit of leaf area, Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, Thickness = leaf thickness, Vessel size = stem conduit size, Vessel density = stem conduit density, WD = wood density. Results shown are from the entire dataset (see methods for details).}
\label{fig:fig1}
\end{figure}

>>>>

Using the identified boundaries among stages we were able to assign all studies into one of three classes: ``seedling'', ``sapling'', or ``adult''. The group ``juvenile'' was largely overlapping with the three others, and so we merged this category with one or the other on a case by case basis. Using the above cut-offs, we were also able to assign a stage category to studies for which no stage was distinctly mentioned by the authors (5\% of the articles, noted as ``mix'' in Fig. \ref{fig:fig2}). Only 6 of 103 articles were ambiguous on their ontogenetic stage attribution. Half of these equivocal cases involved the juvenile category. We attributed a study to a new stage category only when 95\% of the recorded range of sizes for that study lay within a particular stage category. Other equivocal cases were caused by ambiguity in the usage of ``adult'' by the authors. Some plants with a large stature were not described as adult by authors because they were still non reproductive \citep{King:2006he}. Conversely plants with small stature -- because stunted by wind for example \citep{Stratton:2001ck} -- were sometimes assessed as adult when reproductive. For purposes of the present paper, we attributed a study to the adult stage when the authors identified the plants as reproductive, and to sapling stage when they identified the plants as non-reproductive, regardless of the plant's size or age. We found that results using the original categories established by the authors were similar to those using the standardized categories (see Fig. \ref{fig:figA6}). Thus in the main text, we report only results using the standardized categories.

@@

The dataset included studies differing in the growth measurement used (relative or absolute), the number of species, the experimental conditions (stressed vs unstressed), and the concordance between stages used to measure trait and growth rate. These differences provide good criteria for assessing the quality of the data for addressing our hypotheses. We adopted a three-stage approach where at first we were conservative by focusing exclusively on relative growth rate, across studies that included \underline{\textgreater} 10 species, under unstressed conditions, and where the trait and growth were measured at the same stage (except for maximum height and seed mass, where a particular stage is by definition required to measure the trait). We then compared these results with those from the subset of the dataset using relative growth rate, but with all other criteria relaxed. Finally, we compared results from both relative and absolute growth rates, from the entire dataset.

@@

For wood density and seed mass, the available results are consistent with our theoretical expectations, but it would be desirable to have substantially more data. Indeed, while it is possible to conduct a meta-analysis even with few observations, strong conclusions are generally precluded by very large CIs of the effect size (i.e highly uncertain estimates and CI often spanned zero). 
For wood density, results were consistent with the prediction for negative correlations with relative growth rate across all plant sizes (Fig. \ref{fig:fig3}b). Accordingly, model fits were marginally improved when stage was incorporated (p-value = 0.049). Note that the CIs did not cross zero for seedling and sapling stage, despite the small sample sizes (n= 1 and 4 respectively), meaning a strong correlation and/or a good consistent effect in the coefficient reported for each stage across studies. 
As expected, for seed mass the correlation with relative growth rate shifted from negative at seedling stage ($\bar{r}= -0.5 \pm 0.07$) to not significantly different from zero at adult and sapling stages (Fig \ref{fig:fig3}c, CIs overlapping zero). However, the correlations between seed mass and relative growth rate at adult and sapling stages were also not significantly different from that observed at seedling stage (see log likelihood ratio tests and CIs overlapping in Fig. \ref{fig:fig3}c). We could not distinguish between these alternatives. But the number of coefficients reported for seed mass was almost similar between adult (n= 7) and seedling (n= 10) stages, suggesting that CIs overlapping zero at adult stage were due to variation in correlation coefficients reported rather than to a much smaller numbers of replicates.

For maximum height and Aarea, correlations with relative growth rate were also consistent with expectations, but our ability to test all hypotheses was at times limited due to low or no sample size. For maximum height, the expectation was for a correlation with growth to be present for large plants but absent for smaller sizes, the mechanism being increased relative allocation to reproduction as maximum height is approached. Here, maximum height was positively correlated with growth at adult stage ($\bar{r}= 0.38 \pm 0.05$), not significantly correlated at sapling stage (Fig \ref{fig:fig3}c, CIs overlapping zero) and no data were recorded for seedling stage. While these results were consistent with our expectation that maximum height would be correlated with growth at adult stage, they did not differ significantly across ontogenetic stages (model fit not improved when stage was incorporated, see log likelihood ratio tests in Fig. \ref{fig:fig3}c, and CIs overlapping between sapling and adult stages). Due to small numbers of replicate studies for sapling stage and no replicates for seedling stage, we could not conclude on the absence of correlation between maximum height and growth at small plant size.
For Aarea, results were consistent with a positive correlation with relative growth rate at seedling stage. However, no conclusion can be drawn about a change along ontogenetic stages, as only four correlation coefficients were reported for this trait, and three of these were measured at the seedling stage (Fig. \ref{fig:fig3}e).

@@

For five prominent traits, the patterns observed were consistent with specific predictions from a mechanistic growth model, although there were some limitations arising from unequal numbers of observation across ontogenetic stages. This suggests the mechanisms through which traits affect growth are correctly represented in the model. For SLA, the meta-analysis strongly confirmed the predicted influence of plant size on the trait-growth correlation. For wood density, Aarea, seed mass and maximum height results were consistent with predictions, but data was limited and additional studies would be helpful for these trait-growth-stage relationships.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=15cm,height=20cm,keepaspectratio]{output/Fig1}
\caption{Number of recorded correlations between trait and growth rate in our dataset, by stage and trait. Traits are: Aarea (or Amass) = rate of CO$_{2}$ assimilation per unit leaf area (or leaf mass) = maximum photosynthetic rate, Hmax = asymptotic or maximum height, Ks = stem hydraulic conductance, LA = leaf area, LMR = leaf mass ratio, NARarea  (or NARmass) = net mass assimilation rate per unit of leaf area (or leaf mass), Nmass (or Narea) = nitrogen content per unit of leaf mass (or leaf area), Pmass = phosphorus content per unit of leaf mass, SA/LA = sapwood area per leaf area, SLA = specific leaf area, Thickness = leaf thickness, Vessel size = stem conduit size, Vessel density = stem conduit density, WD = wood density. Results shown are from the entire dataset (see methods for details).}
\label{fig:fig1}
\end{figure}
